### 2012-09-28 paper_sean/conclusion.tex
big commit before meeting Jeff/Jason today. Changes: - Shorten - Citation problems - Figure updates (fonts larger) - Figure caption updates - Table right justify - Table sigfigs - Table specified whether max or avg - Adding Kyrre's data to authorlist

% !TEX root =main.tex
HyperNEAT has proven to be an effective encoding in gait evolution with an ability to find gait regularities without human intervention \cite{clune1,clune2,yos:clune}. 
Glette et al. \cite{glette} has shown that evolving gaits in simulation using a simple algorithm with a large number of evaluations could outperform gait evolution using a more sophisticated algorithm in hardware thanks to the simulator affording a larger number of evaluations to be performed for evolution. 
In this study, we took the two -- HyperNEAT and a physical simulator -- and showed that HyperNEAT, when it could afford many evaluations, could evolve impressive gaits. 
In simulation, HyperNEAT gaits outperformed Genetic Algorithm gaits \cite{glette} by 54.3\% and by 5.4\% in the real life trials. 
This study also showed that HyperNEAT will exploit the characteristics of the problem, such as ferquency and regularity, in order to maximize its fitness.
%Without frequency limiters, it would make the legs move really fast in order to increase fitness. 
%When that was no longer allowed, it pushed the servo positions to their boundaries by increasing the stride lengths as wide as possible. 
However, because of the physical limitations of the robot, these exploitations did not transfer well to reality. %could not be transferred well to reality.
Moreover, the inevitable faults with the simulator (it cannot model the robot or the real world perfectly) added to the difficulty in the transfer from simulation to reality. 
Thus, many of the gaits performed much worse than it did in simulation and had difficulty in achieving consistent results.

Next steps for this research heads towards improving the transfer process of the gaits from simulation to reality. 
The performance gap between simulation and reality is called the \emph{reality gap} \cite{jakobi}. 
Many different approaches to reduce the reality gap has been proposed \cite{koos1,bongard,zagal}. 
To improve upon the results from this study, one or more of these approaches could be used to aid in the transfer to reality. 
The one easiest and most promising to implement is the Transferability Approach \cite{koos2} which has shown to be successful in quadrapedal gait evolution. 
HyperNEAT, used with a simulator and a transfer approach, shows much hope in producing yet a better, more coordinated, and higher performing gait on the QuadraTot platform.

>>>>

% !TEX root =main.tex
HyperNEAT has proven to be an effective encoding in gait evolution with an ability to find gait regularities without human intervention~\cite{clune2009evolving,clune2011performance,yos:clune}. Glette et al. 2012~\cite{glette} showed the benefits of a physical simulator by evolving, with a simple genetic algorithm, a gait that outperformed those HyperNEAT evolved in hardware.
In this study, we took the two -- HyperNEAT and a physical simulator -- and showed that HyperNEAT's performance significantly increases with longer evolution periods. The gaits produced in this study outperformed the previous best gait on the QuadraTot by 54.3\% in simulation and by 5.4\% in real life trials. This study also showed that HyperNEAT will exploit the characteristics of the problem, such as frequency and regularity, in order to maximize its fitness.
%In simulation, HyperNEAT gaits outperformed Genetic Algorithm gaits~\cite{glette} by 54.3\% and by 5.4\% in the real life trials.

%Without frequency limiters, it would make the legs move really fast in order to increase fitness. 
%When that was no longer allowed, it pushed the servo positions to their boundaries by increasing the stride lengths as wide as possible. 
However, we faced a problem in the transfer from simulation to reality. %because of the physical limitations of the robot, these exploitations did not transfer well to reality. %could not be transferred well to reality.
One of the reasons was that the robot was greatly underpowered and could not perform the transferred gaits. The other reasons was the inevitable faults with the simulator (as it cannot model the robot or the real world perfectly). Because of these factors, many gaits performed much worse than it did in simulation and had difficulty in achieving consistent results.

Next steps for this research heads towards improving the transfer process of the gaits from simulation to reality by reducing the \emph{reality gap}~\cite{jakobi}. Many different approaches to reduce the reality gap has been proposed~\cite{koos2010crossing,bongard,zagal}. One of the easiest and most promising of these methods is the Transferability Approach~\cite{koos2011transferability}, which has already proven to be successful in quadrapedal gait evolution. Another method that we could take is to evolve the simulator itself to adapt itself to the real world. HyperNEAT, used with a simulator and a transfer approach, shows much hope in producing yet a better, more coordinated, and higher performing gait for many robotic platforms.

### 2012-09-28 paper_sean/hyperneat.tex
big commit before meeting Jeff/Jason today. Changes: - Shorten - Citation problems - Figure updates (fonts larger) - Figure caption updates - Table right justify - Table sigfigs - Table specified whether max or avg - Adding Kyrre's data to authorlist

\figp{hyperneat_bug_example}{.8}{Compositional Pattern Producing Networks (CPPNs). CPPN  genomes use mathematical functions to generate regularities such as symmetry and repetitions. CPPN genomes allow functions to be made of other functions, allowing multiple regularity motifs to be present. Figure taken from Stanley \cite{stanley2007CPPN}.}

@@

In HyperNEAT, the CPPN genomes evolve via the Neuroevolution of Augmenting Topologies (NEAT) algorithm \cite{stanley2006exploiting}. 
The NEAT algorithm has three major components \cite{stanley2006exploiting}. 
First, it starts with small genomes that encode simple networks and complexifies them by mutations that add nodes and links to the network. Second, NEAT has a fitness-sharing system that preserves diversity and allows for new innovations to be tuned by evolution before competing them against more adapted rivals. 
Third, historical information is recorded that facilitates crossover in a way that is effective, yet avoids the need for expensive topological analysis. A full explanation of NEAT can be found in Stanley and Miikkulainen \cite{stanley2006exploiting}.

In this study, the ANN inputs, outputs, and the size of the hidden layer are the same as in Yosinski et al. 2011 \cite{yos:clune}. 
The ANN has a fixed topology of three \begin{math}3\times4\end{math} Cartesian grids of nodes for the input, hidden, and output layers. 
The inputs to the substrate were the angles requested in the prevous time step for each of the 9 joints of the robot and a sine and cosine wave to facilitate periodic motion. 
The outputs of the substrate at each time step were nine numbers (for each joint) in the range [-1, 1] which were scaled to the range [0, 1023], the allowable ranges for the servos. 
As in Yosinski et al. \cite{yos:clune}, to obtain each commanded position we averaged over four outputs of the network to reduce the possibility of switching from extreme negative to extreme positive numbers at a very high frequency.

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=hyperneat.eps, width=8cm}
% \caption [ ]{Producing ANNs from CPPNs. In HyperNEAT, weights are specified as a function of the Cartesian coordinates for source and input nodes and a constant bias for each connection between the source and input. All pairwise combinations of source and target node coordinates are iteratively passed into CPPNs to determine the weight of each ANN link. Figure taken from Clune et al. \cite{clune2}}.
% \end{center}
% \end{figure}
\figp{hyperneatExplanation}{1}{A CPPN specifying a neural network. In HyperNEAT, weights are a function of the Cartesian coordinates of the source and target node for each connection. All pairwise combinations of source and target node coordinates are iteratively passed into a CPPN to determine the weight of each ANN link. Figure from Clune et al. \cite{clune2011performance}.}

>>>>

\figp{hyperneat_bug_example}{.8}{Compositional Pattern Producing Networks (CPPNs). CPPN  genomes use mathematical functions to generate regularities such as symmetry and repetitions. CPPN genomes allow functions to be made of other functions, allowing multiple regularity motifs to be present. Figure taken from Stanley 2007~\cite{stanley2007CPPN}.}

@@

In HyperNEAT, the CPPN genomes evolve via the Neuroevolution of Augmenting Topologies (NEAT) algorithm~\cite{stanley2006exploiting}. 
The NEAT algorithm has three major components~\cite{stanley2006exploiting}. 
First, it starts with small genomes that encode simple networks and complexifies them by mutations that add nodes and links to the network. Second, NEAT has a fitness-sharing system that preserves diversity and allows for new innovations to be tuned by evolution before competing them against more adapted rivals. 
Third, historical information is recorded that facilitates crossover in a way that is effective, yet avoids the need for expensive topological analysis. A full explanation of NEAT can be found in Stanley and Miikkulainen 2006~\cite{stanley2006exploiting}.

In this study, the ANN inputs, outputs, and the size of the hidden layer are the same as in Yosinski et al. 2011~\cite{yos:clune}. 
The ANN has a fixed topology of three \begin{math}3\times4\end{math} Cartesian grids of nodes for the input, hidden, and output layers. 
The inputs to the substrate were the angles requested in the prevous time step for each of the 9 joints of the robot and a sine and cosine wave to facilitate periodic motion. 
The outputs of the substrate at each time step were nine numbers (for each joint) in the range [-1, 1] which were scaled to the range [0, 1023], the allowable ranges for the servos. 
As in Yosinski et al. 2011~\cite{yos:clune}, to obtain each commanded position we averaged over four outputs of the network to reduce the possibility of switching from extreme negative to extreme positive numbers at a very high frequency.

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=hyperneat.eps, width=8cm}
% \caption [ ]{Producing ANNs from CPPNs. In HyperNEAT, weights are specified as a function of the Cartesian coordinates for source and input nodes and a constant bias for each connection between the source and input. All pairwise combinations of source and target node coordinates are iteratively passed into CPPNs to determine the weight of each ANN link. Figure taken from Clune et al.~\cite{clune2}}.
% \end{center}
% \end{figure}
\figp{hyperneatExplanation}{1}{A CPPN specifying a neural network. In HyperNEAT, weights are a function of the Cartesian coordinates of the source and target node for each connection. All pairwise combinations of source and target node coordinates are iteratively passed into a CPPN to determine the weight of each ANN link. Figure from Clune et al. 2011~\cite{clune2011performance}.}

### 2012-09-28 paper_sean/introduction.tex
big commit before meeting Jeff/Jason today. Changes: - Shorten - Citation problems - Figure updates (fonts larger) - Figure caption updates - Table right justify - Table sigfigs - Table specified whether max or avg - Adding Kyrre's data to authorlist

% !TEX root =main.tex
\figgp{quadratotWhiteBooties}{.45}{quadratot_simulator}{.45}{The QuadraTot robot platform on which gaits were evolved. \textbf{Left: } The physical robot, which is composed of 3-D printable and off-the-shelf components.
\textbf{Right: } The robot in the simulator. Figure adapted from from Glette et al. \cite{glette}.}

Robots that can move around on legs can operate in a much wider range of environments than their wheeled counterparts. A downside is that designing gaits for legged robots is a difficult and time-consuming process for human engineers~\cite{strom1999legged,wettergreen1992gait}, and must be repeated every time a new robot is created or the design of an existing robot is changed~\cite{hornby2005autonomous}. 
Because of these difficulties in hand-designing gaits, scientists have investigated automatically producing gaits with machine learning or evolutionary algorithms, an approach that often produces better gaits than those created by human engineers~\cite{valsalam:mii,kohl:stone,hornby2005autonomous,hornby2003generative,yos:clune}. It has often been found that both machine-created and human-created gaits perform better if they are \emph{regular}, in that they have coordinated movements, such as left-right symmetry or front-back symmetry~\cite{valsalam:mii,clune2011performance,clune2009evolving,clune2009sensitivity}.
However, in most previous works, the experimenter has had to manually decide the regularities of the evolved gaits \cite{valsalam:mii,tellez,beer,raibert}. 
Such manual intervention is time consuming, requires expert knowledge, and adds constraints that may hurt performance.

Previous work has shown that the Hypercube-based NeuroEvolution of Augmenting Topologies (HyperNEAT) generative encoding \cite{stanley2009hypercube} can automatically generate a variety of regular gaits and that it outperforms direct encoding controls on this task \cite{clune2009evolving,clune2011performance}.
However, that work only verified these claims in simulation. 
A follow-up paper by Yosinski et al. evolved gaits with HyperNEAT directly in hardware on the QuadraTot robot platform and found that HyperNEAT's gaits outperformed manually designed parameterized learning algorithms, but still did not produce impressive, natural gaits \cite{yos:clune}. 
A further follow-up study built a simulator for the QuadraTot robot platform to test whether the inclusion of a simulator would improve results and found that it did:  when gaits were evolved in this simulator with a simple direct encoding and then transferred to the QuadraTot robot, the resulting gaits were faster than the gaits produced by evolving gaits with HyperNEAT directly on the robot \cite{glette}. 
The simulator helped because it afforded much larger population sizes and more generations than were possible when evolving directly in hardware. Specifically, the simulator allowed 333 times as many evaluations per run as HyperNEAT had in hardware (60000 vs. 180)~\cite{yos:clune,glette}.

The work with the simulator (Glette et al.~\cite{glette}) evolved gaits with a simple encoding that was manually constrained to produce specific regularities. The success of that work raises the question of whether the performance gains were due to the use of a simulator, or the use of a simple, manually-constrained encoding. We hypothesized that HyperNEAT, which has been previously shown to automatically discover complex regularities to produce high-performing gaits~\cite{clune2011performance,yos:clune}, would outperform the simpler encoding from Glette et al.~\cite{glette} if HyperNEAT is combined with a simulator. 
Here we test that hypothesis by evolving gaits with HyperNEAT in the same simulator from Glette et al. and then transferring those gaits to the QuadraTot robot.  Our experiments confirm the hypothesis: gaits evolved with HyperNEAT and then transferred to reality were the highest performing observed to date on the QuadraTot robot platform.

>>>>

% !TEX root =main.tex
\figgp{quadratotWhiteBooties}{.45}{quadratot_simulator}{.45}{The QuadraTot robot platform on which gaits were evolved. \textbf{Left: } The physical robot, which is composed of 3-D printable and off-the-shelf components.
\textbf{Right: } The robot in the simulator. Figure adapted from from Glette et al. 2012~\cite{glette}.}

Robots that can move around on legs can operate in a much wider range of environments than their wheeled counterparts. A downside is that designing gaits for legged robots is a difficult and time-consuming process for human engineers~\cite{strom1999legged,wettergreen1992gait}, and must be repeated every time a new robot is created or the design of an existing robot is changed~\cite{hornby2005autonomous}. 
Because of these difficulties in hand-designing gaits, scientists have investigated automatically producing gaits with machine learning or evolutionary algorithms, an approach that often produces better gaits than those created by human engineers~\cite{valsalam:mii,kohl:stone,hornby2005autonomous,hornby2003generative,yos:clune}. It has often been found that both machine-created and human-created gaits perform better if they are \emph{regular}, in that they have coordinated movements, such as left-right symmetry or front-back symmetry~\cite{valsalam:mii,clune2011performance,clune2009evolving,clune2009sensitivity}.
However, in most previous works, the experimenter has had to manually decide the regularities of the evolved gaits~\cite{valsalam:mii,tellez,beer,raibert}. 
Such manual intervention is time consuming, requires expert knowledge, and adds constraints that may hurt performance.

Previous work has shown that the Hypercube-based NeuroEvolution of Augmenting Topologies (HyperNEAT) generative encoding~\cite{stanley2009hypercube} can automatically generate a variety of regular gaits and that it outperforms direct encoding controls on this task~\cite{clune2009evolving,clune2011performance}.
However, that work only verified these claims in simulation. 
A follow-up paper by Yosinski et al. evolved gaits with HyperNEAT directly in hardware on the QuadraTot robot platform and found that HyperNEAT's gaits outperformed manually designed parameterized learning algorithms, but still did not produce impressive, natural gaits~\cite{yos:clune}. 
A further follow-up study built a simulator for the QuadraTot robot platform to test whether the inclusion of a simulator would improve results and found that it did:  when gaits were evolved in this simulator with a simple direct encoding and then transferred to the QuadraTot robot, the resulting gaits were faster than the gaits produced by evolving gaits with HyperNEAT directly on the robot~\cite{glette}. 
The simulator helped because it afforded much larger population sizes and more generations than were possible when evolving directly in hardware. Specifically, the simulator allowed 333 times as many evaluations per run as HyperNEAT had in hardware (60000 vs. 180)~\cite{yos:clune,glette}.

The work with the simulator (Glette et al. 2012~\cite{glette}) evolved gaits with a simple encoding that was manually constrained to produce specific regularities. The success of that work raises the question of whether the performance gains were due to the use of a simulator, or the use of a simple, manually-constrained encoding. We hypothesized that HyperNEAT, which has been previously shown to automatically discover complex regularities to produce high-performing gaits~\cite{clune2011performance,yos:clune}, would outperform the simpler encoding from Glette et al. 2012~\cite{glette} if HyperNEAT is combined with a simulator. 
Here we test that hypothesis by evolving gaits with HyperNEAT in the same simulator from Glette et al. and then transferring those gaits to the QuadraTot robot.  Our experiments confirm the hypothesis: gaits evolved with HyperNEAT and then transferred to reality were the highest performing observed to date on the QuadraTot robot platform.

### 2012-09-28 paper_sean/results.tex
big commit before meeting Jeff/Jason today. Changes: - Shorten - Citation problems - Figure updates (fonts larger) - Figure caption updates - Table right justify - Table sigfigs - Table specified whether max or avg - Adding Kyrre's data to authorlist

% !TEX root =main.tex
Because of the noise inherent in all physical robots, especially different copies of robots, the best way to compare two algorithms is in simulation. The gaits evolved with HyperNEAT in simulation significantly outperformed the gaits evolved with a GA by Glette et al. 2012~(Table~\ref{resultsTable}). INSERT-RES-OF-SIM-PARAGRAPH.

@@

% table!
\begin{table}
\caption{The number of evaluations per run (gen x orgs), simulated and real trial velocities in [cm/s]. Data used from Yosinksi et al. \cite{yos:clune}, Glette et al. \cite{glette}, and Shen et al. \cite{haocheng}.
*17.8 cm/s was recorded on Glette et al. \cite{glette} but the same gait tried on our robot averaged over ten runs was 12.951 cm/s}  %\tablabel{results}
\begin{center}
\begin{tabular}{|l|c|c|c|c|}
\hline
                                         & Evaluations  & Simulated Velocity  & Real Velocity \\
\hline
Best parameter optimization learning    &153    & --    & 4.64 \\
\hline
Hardware HyperNEAT                  & 180         & --         &   9.7     \\
\hline
Simulated GA              & 60000       & 16.4       &   12.951*     \\
\hline
RL PoWER Spline                          & 300         & --         &   11.05 \\
\hline
Simulated HyperNEAT                      & 40000       & 25.4       &   14.5 \\
\hline
\end{tabular}
\end{center}
\label{resultsTable}
\end{table}
%JMC All numbers in the same column should be to the same precision (e.g. for the real velocity column, one decimal place). Also, can you right-justify the numbers (while still keeping them roughly centered)? Comparing numbers that are not right-justified is hard...so if you can't center and right-justify, then I let's see if just having it be right justified is better.

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=fit_gen.eps, width=8cm}
% \caption[ ]{Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%.}
% \end{center}
% \end{figure}
\figgp{avg_fit_vs_gen}{.48}{avg_freq_vs_gen}{.48}{\textbf{Left: }Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%. \textbf{Right: }Average frequencies of the gaits averaged over 20 runs. After implementing a policy which punished gaits with high frequencies, HyperNEAT was able to find high-performing, low frequency gaits. This policy was implemented to reduce servo shutdowns which heavily affected the reality performance of the robot.}

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=servo_pos.eps, width=8cm}
% \caption[ ]{Top: A plot of servo positions for a gait produced in this study. Bottom: A plot of servo positions for a gait produced by HyperNEAT in hardware. Figure taken from Yosinski et al. \cite{yos:clune}. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}
% \end{center}
% \end{figure}
%\figgp{servo_plot_111}{.6}{neat_110115_211410_00000_002_filt_zoom}{.6}{Top: A plot of servo positions for a gait produced in this study. Bottom: A plot of servo positions for a gait produced by HyperNEAT in hardware. Figure taken from Yosinski et al. \cite{yos:clune}. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}%\edit{wrong fig on top, regenerate top fig as pdf and add.}}
\figp{servo_plot_111}{.6}{A plot of servo positions for a gait produced in this study. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}%\edit{wrong fig on top, regenerate top fig as pdf and add.}}

>>>>

% !TEX root =main.tex
Because of the noise inherent in all physical robots, especially with different copies of robots, the best way to compare two algorithms is in simulation. The gaits evolved with HyperNEAT in simulation significantly outperformed the gaits evolved with a GA by Glette et al. 2012~(Table~\ref{resultsTable}). INSERT-RES-OF-SIM-PARAGRAPH.

@@

% table!
\begin{table}
\caption{The number of evaluations per run, simulated (average) and real trial (max) velocities in [cm/s]. Data used from Yosinksi et al. 2011~\cite{yos:clune}, Glette et al. 2012~\cite{glette}, and Shen et al. 2012~\cite{haocheng}.
*17.8 cm/s was recorded on Glette et al. 2012~\cite{glette} but the same gait tried on our robot averaged over ten runs was 12.951 cm/s}  %\tablabel{results}
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline
                                         & Evaluations  & Simulated Velocity  & Real Velocity \\
\hline
Best parameter optimization learning    &153    & --    & 4.6 \\
\hline
Hardware HyperNEAT                  & 180         & --         &   9.7     \\
\hline
Simulated GA              & 60000       & 16.4       &   13.0*     \\
\hline
RL PoWER Spline                          & 300         & --         &   11.1 \\
\hline
Simulated HyperNEAT                      & 40000       & 25.4       &   14.5 \\
\hline
\end{tabular}
\end{center}
\label{resultsTable}
\end{table}
%JMC All numbers in the same column should be to the same precision (e.g. for the real velocity column, one decimal place). Also, can you right-justify the numbers (while still keeping them roughly centered)? Comparing numbers that are not right-justified is hard...so if you can't center and right-justify, then I let's see if just having it be right justified is better.

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=fit_gen.eps, width=8cm}
% \caption[ ]{Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%.}
% \end{center}
% \end{figure}
\figgp{avg_fit_vs_gen}{.48}{avg_freq_vs_gen}{.48}{Performance plots. \textbf{Left: }Fitnesses of the champion gaits averaged over 20 runs, measured in distance travelled over 12 seconds [cm/s]. These gaits outperformed gaits evolved in simulation using Genetic Algorithm by 54.3\% in simulation and 5.4\% better in reality while outperforming gaits evolved using HyperNEAT in hardware by 49.5\%. \textbf{Right: }Average frequencies of the gaits averaged over 20 runs. After implementing a policy which punished gaits with high frequencies, HyperNEAT was able to find high-performing, low frequency gaits. This policy was implemented to reduce servo shutdowns which heavily affected the reality performance of the robot.}

@@

% \begin{figure}
% \begin{center}
% \vspace{1cm}
% \epsfig{file=servo_pos.eps, width=8cm}
% \caption[ ]{Top: A plot of servo positions for a gait produced in this study. Bottom: A plot of servo positions for a gait produced by HyperNEAT in hardware. Figure taken from Yosinski et al.~\cite{yos:clune}. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}
% \end{center}
% \end{figure}
%\figgp{servo_plot_111}{.6}{neat_110115_211410_00000_002_filt_zoom}{.6}{Top: A plot of servo positions for a gait produced in this study. Bottom: A plot of servo positions for a gait produced by HyperNEAT in hardware. Figure taken from Yosinski et al.~\cite{yos:clune}. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}%\edit{wrong fig on top, regenerate top fig as pdf and add.}}
\figp{servo_plot_111}{0.95}{A plot of servo positions for a gait produced in this study. The smoothness, symmetry, and regularities of the top plot shows that by using a simulator, HyperNEAT was able to fully exploit the geometry of the problem.}

### 2012-09-28 paper_sean/robot.tex
big commit before meeting Jeff/Jason today. Changes: - Shorten - Citation problems - Figure updates (fonts larger) - Figure caption updates - Table right justify - Table sigfigs - Table specified whether max or avg - Adding Kyrre's data to authorlist

We performed the experiments on the QuadraTot quadrupedal robot platform~\cite{yos:clune}.
It has 9 degrees-of-freedom (DOF): two joints per leg and one joint that allowed rotation along the robot's midline. %we def want a picture of the cool robot. I propose putting the picture of the robot on the left and the picture of the simulated robot on the right of the same figure. That also does a nice job of highlighting the theme of the paper (robot+sim) in one figure. Try to get this figure to show up as early as possible in the paper (e.g top of the 2nd page). 
QuadraTot has open source software and hardware\footnote{A list of all of the parts and all software, including the simulator, is available for free at http://quadratot.yosinski.com.} %JMC: This should really be a Creative Machines/Cornell web address. Can you ask Jason if he agrees and, if so, take the lead on making sure it gets made?
, including many parts that can be 3D printed. These traits enable other labs to perform research with it, providing a platform for comparing different gait-learning algorithms. 
To date there are results on the platform from nine different learning algorithms from three previous publications \cite{yos:clune,glette,haocheng}. We identified the robot's position using infrared LEDs on the robot which was tracked using a Wiimote~\cite{yos:clune}.

Both in this work and that of Glette et al.~\cite{glette}, the joints are powered by Robotis Dynamixel servos; five AX-18A servos for the inner joint of each leg and the single midline joint, and four AX-12A servos for the the outer joint of each leg, which requires less power and can thus have less expensive motors. The servos were sent new positions at 40Hz via the Pydynamixel library.

Each servo has a built-in safety mechanism that shuts itself off to prevent damage if the servo's current, range, temperature, or torque was too high \cite{robotis}. During evolution, this safety mechanism frequently activated, and did so inconsistently, adding significant noise to the evaluation process. As pointed out in Yosinski et al. \cite{yos:clune}, gaits generated during evolution are highly variable and generate many shutdowns because they forced the servos to exert too much torque or command positions that are out of range. %JMC: Does requesting a value out of range really cause the servo to shut down? That seems like something we could easily fix (by never allowing an out-of-range value). Now that we do not allow values out of range, we still see lots of shutdowns...making me think that the problem was really torque. What do you think?
To minimize shutdowns, we limited the allowable range of movement for the inner, outer, and hip joints to [-85\degree, +60\degree], [-113\degree, +39\degree], and [-28\degree, +28\degree], respectively. We also implemented the Smart Cropping System from Shen et al. \cite{haocheng}, which prevents combinations of joint positions for the inner and outer joint of each leg that generate extreme amounts of torque. A final method of reducing torque was to reduce the weight of the robot. Yosinski et al.~\cite{yos:clune} had the small Linux computer that performed all computation on the robot, but we removed it and sent commands to the robot via a cable.

>>>>

We performed the experiments on the QuadraTot quadrupedal robot platform~\cite{yos:clune}.
It has 9 degrees-of-freedom (DOF): two joints per leg and one joint that allowed rotation along the robot's midline. %we def want a picture of the cool robot. I propose putting the picture of the robot on the left and the picture of the simulated robot on the right of the same figure. That also does a nice job of highlighting the theme of the paper (robot+sim) in one figure. Try to get this figure to show up as early as possible in the paper (e.g top of the 2nd page). 
QuadraTot has open source software and hardware\footnote{A list of all of the parts and all software, including the simulator, is available for free at http://quadratot.yosinski.com.} %JMC: This should really be a Creative Machines/Cornell web address. Can you ask Jason if he agrees and, if so, take the lead on making sure it gets made?
, including many parts that can be 3D printed. These traits enable other labs to perform research with it, providing a platform for comparing different gait-learning algorithms. 
To date there are results on the platform from nine different learning algorithms from three previous publications~\cite{yos:clune,glette,haocheng}. We identified the robot's position using infrared LEDs on the robot which was tracked using a Wiimote~\cite{yos:clune}.

Both in this work and that of Glette et al. 2012~\cite{glette}, the joints are powered by Robotis Dynamixel servos; five AX-18A servos for the inner joint of each leg and the single midline joint, and four AX-12A servos for the the outer joint of each leg, which requires less power and can thus have less expensive motors. The servos were sent new positions at 40Hz via the Pydynamixel library.

Each servo has a built-in safety mechanism that shuts itself off to prevent damage if the servo's current, range, temperature, or torque was too high~\cite{robotis}. During evolution, this safety mechanism frequently activated, and did so inconsistently, adding significant noise to the evaluation process. As pointed out in Yosinski et al. 2011~\cite{yos:clune}, gaits generated during evolution are highly variable and generate many shutdowns because they forced the servos to exert too much torque or command positions that are out of range. %JMC: Does requesting a value out of range really cause the servo to shut down? That seems like something we could easily fix (by never allowing an out-of-range value). Now that we do not allow values out of range, we still see lots of shutdowns...making me think that the problem was really torque. What do you think?
To minimize shutdowns, we limited the allowable range of movement for the inner, outer, and hip joints to [-85\degree, +60\degree], [-113\degree, +39\degree], and [-28\degree, +28\degree], respectively. We also implemented the Smart Cropping System from Shen et al. 2012~\cite{haocheng}, which prevents combinations of joint positions for the inner and outer joint of each leg that generate extreme amounts of torque. A final method of reducing torque was to reduce the weight of the robot. Yosinski et al. 2011~\cite{yos:clune} had the small Linux computer that performed all computation on the robot, but we removed it and sent commands to the robot via a cable.
